Machine learning high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2023-03-13 qcd.tex
missing + signs in dipole formulae
improve spacing

    \subsection{Final-initial dipoles}
        For the case of final-state emitter and initial-state
        spectator we have the dipole factorisation formula
        \begin{equation}\label{eqn:FI_factorisation}
            \begin{split}
                {}_{n+1}\langle 1, \ldots, n+1; a, \ldots |& 1, \ldots, n+1; a, \ldots \rangle_{n+1} = \\
                &\sum_{i,j}\sum_{k\neq i,j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},\ldots) \\
                &\sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},\ldots) + \ldots \, ,
            \end{split}
        \end{equation}
        where the new dipole contribution appears on the second line.
        The initial-state parton is denoted by $a$.
        The dipole term $\mathcal{D}_{ij}^{a}$ is given by
        \begin{equation}\label{eqn:D_aij}
            \begin{split}
                \mathcal{D}_{ij}^{a}(p_{1}, \ldots, p_{n+1}; p_{a}, \ldots) &= -\dfrac{1}{2p_{i}p_{j}} \dfrac{1}{x_{ij,a}} \\
                {}_{n} \langle 1, \ldots, \widetilde{ij}, &\ldots, n+1; \tilde{a}, \ldots | \dfrac{\boldsymbol{T}_{a} \cdot \boldsymbol{T}_{ij}}{\boldsymbol{T}_{ij}^{2}} \boldsymbol{V}_{ij}^{a} | 1, \ldots, \widetilde{ij}, \ldots, n+1; \tilde{a} , \ldots \rangle_{n} \, .
            \end{split}
        \end{equation}
        where $\widetilde{ij}$ is the composite emitter in the
        final-state, and the spectator $\tilde{a}$ is in the initial-state.
        The momentum mapping for $p_{i} + p_{j} + p_{a} \rightarrow \tilde{p}_{ij} + \tilde{p}_{a}$ is given as
        \begin{equation}\label{eqn:FI_mapping}
            \tilde{p}_{ij}^{\mu} = p_{i}^{\mu} + p_{j}^{\mu} - (1-x_{ij,a})p_{a}^{\mu} \, , \quad \tilde{p}_{a}^{\mu} = x_{ij,a} \, p_{a}^{\mu} \, , \\
        \end{equation}
        such that momentum conservation and on-shell conditions are met:
        \begin{equation}\label{eqn:FI_mapping_conditions}
            \begin{split}
                &p_{i}^{\mu} + p_{j}^{\mu} - p_{a}^{\mu} = \tilde{p}_{ij}^{\mu} - \tilde{p}_{a}^{\mu} \, , \\
                &\tilde{p}_{ij}^{2} = \tilde{p}_{a}^{2} = 0 \, .
            \end{split}
        \end{equation}
        The recoil parameter $x_{ij,a}$ and splitting variables $\tilde{z}$ are given by
        \begin{equation}\label{eqn:x_ija}
        \begin{split}
                &x_{ij,a} = \dfrac{p_{i}p_{a} + p_{j}p_{a} - p_{i}p_{j}}{(p_{i}p_{a} + p_{j}p_{a})} \, , \\
                &\tilde{z}_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} = 1 - \tilde{z}_{i} \, .
            \end{split}
        \end{equation}
        The spin-averaged splitting functions are given as
        \begin{align}\label{eqn:FI_V_aij}
            &\langle \boldsymbol{V}_{q_{i}g_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} C_{F} \left[\dfrac{2}{1 - \tilde{z}_{i} + (1-x_{ij,a})} - (1+\tilde{z}_{i}) - \epsilon \tilde{z}_{j} \right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{q_{i}\bar{q}_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} T_{R} \, \left[1-\dfrac{2\tilde{z}_{i}\tilde{z}_{j}}{1-\epsilon}\right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{g_{i}g_{j},a} \rangle = 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{1}{\tilde{z}_{j} + (1-x_{ij,a})} + \dfrac{1}{\tilde{z}_{i} + (1-x_{ij,a})} - 2 + \tilde{z}_{i}\tilde{z}_{j}\right] \, .
        \end{align}

    \subsection{Initial-final dipoles}
        For the case of initial-state emitter and final-state spectator,
        the dipole factorisation formula is given as
        \begin{equation}\label{eqn:IF_factorisation}
            {}_{n+1} \langle 1, \ldots, n+1; a | 1, \ldots, n+1; a \rangle_{n+1} = \sum_{a,i}\sum_{k \neq i}\mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a}) + \ldots \, ,
        \end{equation}
        where there is only one initial-state parton. The dipole is given by
        \begin{equation}\label{eqn:D_aik}
            \begin{split}
            \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a}) &= -\dfrac{1}{2p_{a}p_{i}}\dfrac{1}{x_{ik,a}} \\
            {}_{n}\langle 1, \ldots, \tilde{k}, &\ldots, n+1 ; \widetilde{ai} | \dfrac{\boldsymbol{T}_{k} \cdot \boldsymbol{T}_{ai}}{\boldsymbol{T}_{ai}^{2}} \boldsymbol{V}_{k}^{ai} | 1, \ldots, \tilde{k}, \ldots, n+1; \widetilde{ai} \rangle_{n} \, .
            \end{split}
        \end{equation}
        where the emitter is the initial-state parton $\widetilde{ai}$
        and the spectator is the final-state parton $\tilde{k}$.
        The momenta mapping for $p_{a} + p_{i} + p_{k} \rightarrow \tilde{p}_{ai} + \tilde{p}_{k}$ is
        \begin{equation}\label{eqn:IF_mapping}
            \tilde{p}_{ai}^{\mu} = x_{ik,a} \, p_{a}^{\mu} \, , \quad \tilde{p}_{k}^{\mu} = p_{k}^{\mu} + p_{i}^{\mu} - (1-x_{ik,a})p_{a}^{\mu} \, ,
        \end{equation}
        such that
        \begin{equation}\label{eqn:IF_mapping_conditions}
            \begin{split}
            &p_{i}^{\mu} + p_{k}^{\mu} - p_{a}^{\mu} = \tilde{p}_{k}^{\mu} - \tilde{p}_{ai}^{\mu} \, , \\
            &\tilde{p}_{ai}^{2} = \tilde{p}_{k}^{2} = 0 \, .
            \end{split}
        \end{equation}
        Note that the momentum $\tilde{p}_{ai}$ is parallel
        to $p_{a}$, therefore the modified spectator momentum $\tilde{p}_{k}$
        has to recoil the transverse momentum. This leads to $x_{ik,a}$
        acting as a splitting variable.
        With $x_{ik,a}$ taking the role of $\tilde{z}$, there is another
        parameter, $u_{i}$, appearing in the splitting functions.
        These variables are given as
        \begin{equation}\label{eqn:x_ika}
            x_{ik,a} = \dfrac{p_{k}p_{a}+p_{i}p_{a}-p_{i}p_{k}}{p_{i}p_{a} + p_{k}p_{a}} \, , \quad u_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{k}p_{a}} \, .
        \end{equation}

        The spin-averaged splitting functions are
        \begin{align}\label{eqn:V_aik}
            &\langle \boldsymbol{V}_{k}^{q_{a}g_{i}} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-x_{ik,a} + u_{i}} + (1+x_{ik,a}) - \epsilon(1-x_{ik,a})\right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{k}^{g_{a}\bar{q}_{i}} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[1 - \dfrac{2x_{ik,a}(1-x_{ik,a})}{1-\epsilon}\right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{k}^{g_{a}g_{i}} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{g})} \; 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{1}{1-x_{ik,a} + u_{i}} + \dfrac{1-x_{ik,a}}{x_{ik,a}} - 1 + x_{ik,a}(1-x_{ik,a}) \right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{k}^{q_{a}q_{i}} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{g})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[(1-\epsilon)x_{ik,a} + 2\dfrac{1-x_{ik,a}}{x_{ik,a}} \right] \, ,
        \end{align}
        where $n_{s}(a)$ is the number of polarisations of particle $a$.
        For fermions $n_{s}(q) = n_{s}(\bar{q}) = 2$
        and for gluons $n_{s}(g) = d-2$. In the limit $\epsilon \rightarrow 0$
        all ratios of $n_{s}$ cancel to give unity.

    \subsection{Initial-initial dipoles}
        In the case of two initial-state partons, $a$ and $b$, there
        is an additional dipole contribution compared to the initial-final case.
        The dipole factorisation is given by
        \begin{equation}\label{eqn:II_factorisation}
            \begin{split}
                {}_{n+1} \langle 1, \ldots, n+1; a,b &| 1, \ldots, n+1;a,b \rangle_{n+1} = \\
                &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) \\
                &\sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) + \ldots \, ,
            \end{split}
        \end{equation}
        where the new contribution is the dipole $\mathcal{D}^{ai,b}$ given by
        \begin{equation}\label{eqn:D_aib}
            \begin{split}
                \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) &= -\dfrac{1}{2p_{a}p_{i}} \dfrac{1}{x_{i,ab}} \\
                &{}_{n}\langle \tilde{1}, \ldots, \widetilde{n+1}; \widetilde{ai}, b | \dfrac{\boldsymbol{T}_{b}\cdot\boldsymbol{T}_{ai}}{\boldsymbol{T}^{2}_{ai}}\boldsymbol{V}^{ai,b} | \tilde{1}, \ldots, \widetilde{n+1}; \widetilde{ai}, b \rangle_{n} \, .
            \end{split}
        \end{equation}
        The initial-state parton $\widetilde{ai}$ is the emitter and the
        other initial-state parton $b$ is the spectator. Notice that the
        momenta not involved in the dipole term have been modified with only
        parton $b$ remaining unchanged. The momentum mapping for this case
        has $\tilde{p}_{ai}$ parallel with $p_{a}$ and also
        modifies the momenta for all other final-state momenta (even non-QCD particles) $k_{j}$:
        \begin{align}\label{eqn:II_mapping}
            \tilde{p}_{ai}^{\mu} &= x_{i,ab} \, p_{a}^{\mu} \, , \nonumber \\
            \tilde{k}_{j}^{\mu} &= k_{j}^{\mu} - \dfrac{2k_{j} \cdot (K+\widetilde{K})}{(K+\widetilde{K})^{2}}(K+\widetilde{K})^{\mu} + \dfrac{2k_{j} \cdot K}{K^{2}}\widetilde{K}^{\mu} \, , \\
            x_{i,ab} &= \dfrac{p_{a}p_{b} - p_{i}p_{a}-p_{i}p_{b}}{p_{a}p_{b}} \, . \nonumber
        \end{align}
        where $K$ and $\widetilde{K}$ are the total momenta
        of the dipole before and after the mapping, respectively. They are given by
        \begin{align}\label{eqn:II_mapping2}
            \begin{split}
            K^{\mu} &= p_{a}^{\mu} + p_{b}^{\mu} - p_{i}^{\mu} \, ,\\
            \widetilde{K}^{\mu} &= \tilde{p}_{ai}^{\mu} + p_{b}^{\mu} \, . \\
            \end{split}
        \end{align}
        This mapping conserves momentum and keeps mapped momenta on-shell
        \begin{align}\label{eqn:II_mapping_conditions}
            \begin{split}
            &p_{a}^{\mu} + p_{b}^{\mu} - p_{i}^{\mu} - \sum_{j} k_{j}^{\mu} = \tilde{p}_{ai}^{\mu} + p_{b}^{\mu} - \sum_{j} \tilde{k}_{j}^{\mu} = 0\, , \\
            &\tilde{p}_{ai}^{2} = \tilde{k}_{j}^{2} = 0 \, .
            \end{split}
        \end{align}
        The spin-averaged splitting functions are proportional
        to the Altarelli-Parisi splitting functions which we
        quote here for completeness
        \begin{align}\label{V_aib}
            \begin{split}
            &\langle \boldsymbol{V}^{q_{a}q_{i},b} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{g})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{1 + (1-x_{i,ab})^{2}}{x_{i,ab}}\right] \, , \\
            &\langle \boldsymbol{V}^{q_{a}g_{i},b} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{1 + x_{i,ab}^{2}}{1 - x_{i,ab}}\right] \, , \\
            &\langle \boldsymbol{V}^{g_{a}\bar{q}_{i},b} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[x_{i,ab}^{2} + (1-x_{i,ab})^{2}\right] \, , \\
            &\langle \boldsymbol{V}^{g_{a}g_{i},b} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{g})} \; 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{x_{i,ab}}{1-x_{i,ab}} + \dfrac{1-x_{i,ab}}{x_{i,ab}} + x_{i,ab}(1-x_{i,ab})\right] \, .
            \end{split}
        \end{align}

    \subsection{Master factorisation formula}
        The combination of all types of dipole contributions
        can be combined into a master dipole factorisation formula
        \begin{equation}\label{eqn:dipole_factorisation_master}
            \begin{split}
            {}_{n+1} \langle 1, &\ldots, n+1 | 1, \ldots, n+1 \rangle_{n+1} = \\
            &\sum_{i,j}\sum_{k \neq i, j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) \\
            &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \ldots \, ,
            \end{split}
        \end{equation}
        where it is understood that for a given partonic process,
        the matrix element in the different soft and collinear limits will be
        reproduced by summing over all the relevant dipoles, meaning that
        the dipole functions can be used as a set of functional behaviours for
        approximating the matrix elements for arbitrary SM processes.
        This will be put into practice in Chapters \ref{chapter:fame1}
        and \ref{chapter:fame3}. The terms represented as ellipses in
        the dipole factorisation formulae are non-divergent
        in all of phase-space.

>>>>

    \subsection{Final-initial dipoles}
        For the case of final-state emitter and initial-state
        spectator we have the dipole factorisation formula
        \begin{equation}\label{eqn:FI_factorisation}
            \begin{split}
                {}_{n+1}\langle 1, \ldots, n+1; a, \ldots |& 1, \ldots, n+1; a, \ldots \rangle_{n+1} = \\
                &\sum_{i,j}\sum_{k\neq i,j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},\ldots) \\
                + &\sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},\ldots) + \ldots \, ,
            \end{split}
        \end{equation}
        where the new dipole contribution appears on the second line.
        The initial-state parton is denoted by $a$.
        The dipole term $\mathcal{D}_{ij}^{a}$ is given by
        \begin{equation}\label{eqn:D_aij}
            \begin{split}
                \mathcal{D}_{ij}^{a}(p_{1}, \ldots, p_{n+1}; p_{a}, \ldots) &= -\dfrac{1}{2p_{i}p_{j}} \dfrac{1}{x_{ij,a}} \\
                {}_{n} \langle 1, \ldots, \widetilde{ij}, &\ldots, n+1; \tilde{a}, \ldots | \dfrac{\boldsymbol{T}_{a} \cdot \boldsymbol{T}_{ij}}{\boldsymbol{T}_{ij}^{2}} \boldsymbol{V}_{ij}^{a} | 1, \ldots, \widetilde{ij}, \ldots, n+1; \tilde{a} , \ldots \rangle_{n} \, .
            \end{split}
        \end{equation}
        where $\widetilde{ij}$ is the composite emitter in the
        final-state, and the spectator $\tilde{a}$ is in the initial-state.
        The momentum mapping for $p_{i} + p_{j} + p_{a} \rightarrow \tilde{p}_{ij} + \tilde{p}_{a}$ is given as
        \begin{equation}\label{eqn:FI_mapping}
            \tilde{p}_{ij}^{\mu} = p_{i}^{\mu} + p_{j}^{\mu} - (1-x_{ij,a})p_{a}^{\mu} \, , \quad \tilde{p}_{a}^{\mu} = x_{ij,a} \, p_{a}^{\mu} \, , \\
        \end{equation}
        such that momentum conservation and on-shell conditions are met:
        \begin{equation}\label{eqn:FI_mapping_conditions}
            \begin{split}
                &p_{i}^{\mu} + p_{j}^{\mu} - p_{a}^{\mu} = \tilde{p}_{ij}^{\mu} - \tilde{p}_{a}^{\mu} \, , \\
                &\tilde{p}_{ij}^{2} = \tilde{p}_{a}^{2} = 0 \, .
            \end{split}
        \end{equation}
        The recoil parameter $x_{ij,a}$ and splitting variables $\tilde{z}$ are given by
        \begin{equation}\label{eqn:x_ija}
        \begin{split}
                &x_{ij,a} = \dfrac{p_{i}p_{a} + p_{j}p_{a} - p_{i}p_{j}}{(p_{i}p_{a} + p_{j}p_{a})} \, , \\
                &\tilde{z}_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} = 1 - \tilde{z}_{i} \, .
            \end{split}
        \end{equation}
        The spin-averaged splitting functions are given as
        \begin{align}\label{eqn:FI_V_aij}
            &\langle \boldsymbol{V}_{q_{i}g_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} C_{F} \left[\dfrac{2}{1 - \tilde{z}_{i} + (1-x_{ij,a})} - (1+\tilde{z}_{i}) - \epsilon \tilde{z}_{j} \right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{q_{i}\bar{q}_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} T_{R} \, \left[1-\dfrac{2\tilde{z}_{i}\tilde{z}_{j}}{1-\epsilon}\right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{g_{i}g_{j},a} \rangle = 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{1}{\tilde{z}_{j} + (1-x_{ij,a})} + \dfrac{1}{\tilde{z}_{i} + (1-x_{ij,a})} - 2 + \tilde{z}_{i}\tilde{z}_{j}\right] \, .
        \end{align}

    \subsection{Initial-final dipoles}
        For the case of initial-state emitter and final-state spectator,
        the dipole factorisation formula is given as
        \begin{equation}\label{eqn:IF_factorisation}
            {}_{n+1} \langle 1, \ldots, n+1; a | 1, \ldots, n+1; a \rangle_{n+1} = \sum_{a,i}\sum_{k \neq i}\mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a}) + \ldots \, ,
        \end{equation}
        where there is only one initial-state parton. The dipole is given by
        \begin{equation}\label{eqn:D_aik}
            \begin{split}
            \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a}) &= -\dfrac{1}{2p_{a}p_{i}}\dfrac{1}{x_{ik,a}} \\
            {}_{n}\langle 1, \ldots, \tilde{k}, &\ldots, n+1 ; \widetilde{ai} | \dfrac{\boldsymbol{T}_{k} \cdot \boldsymbol{T}_{ai}}{\boldsymbol{T}_{ai}^{2}} \boldsymbol{V}_{k}^{ai} | 1, \ldots, \tilde{k}, \ldots, n+1; \widetilde{ai} \rangle_{n} \, .
            \end{split}
        \end{equation}
        where the emitter is the initial-state parton $\widetilde{ai}$
        and the spectator is the final-state parton $\tilde{k}$.
        The momenta mapping for $p_{a} + p_{i} + p_{k} \rightarrow \tilde{p}_{ai} + \tilde{p}_{k}$ is
        \begin{equation}\label{eqn:IF_mapping}
            \tilde{p}_{ai}^{\mu} = x_{ik,a} \, p_{a}^{\mu} \, , \quad \tilde{p}_{k}^{\mu} = p_{k}^{\mu} + p_{i}^{\mu} - (1-x_{ik,a})p_{a}^{\mu} \, ,
        \end{equation}
        such that
        \begin{equation}\label{eqn:IF_mapping_conditions}
            \begin{split}
            &p_{i}^{\mu} + p_{k}^{\mu} - p_{a}^{\mu} = \tilde{p}_{k}^{\mu} - \tilde{p}_{ai}^{\mu} \, , \\
            &\tilde{p}_{ai}^{2} = \tilde{p}_{k}^{2} = 0 \, .
            \end{split}
        \end{equation}
        Note that the momentum $\tilde{p}_{ai}$ is parallel
        to $p_{a}$, therefore the modified spectator momentum $\tilde{p}_{k}$
        has to recoil the transverse momentum. This leads to $x_{ik,a}$
        acting as a splitting variable.
        With $x_{ik,a}$ taking the role of $\tilde{z}$, there is another
        parameter, $u_{i}$, appearing in the splitting functions.
        These variables are given as
        \begin{equation}\label{eqn:x_ika}
            x_{ik,a} = \dfrac{p_{k}p_{a}+p_{i}p_{a}-p_{i}p_{k}}{p_{i}p_{a} + p_{k}p_{a}} \, , \quad u_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{k}p_{a}} \, .
        \end{equation}
        The spin-averaged splitting functions are
        \begin{align}\label{eqn:V_aik}
            &\langle \boldsymbol{V}_{k}^{q_{a}g_{i}} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-x_{ik,a} + u_{i}} + (1+x_{ik,a}) - \epsilon(1-x_{ik,a})\right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{k}^{g_{a}\bar{q}_{i}} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[1 - \dfrac{2x_{ik,a}(1-x_{ik,a})}{1-\epsilon}\right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{k}^{g_{a}g_{i}} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{g})} \; 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{1}{1-x_{ik,a} + u_{i}} + \dfrac{1-x_{ik,a}}{x_{ik,a}} - 1 + x_{ik,a}(1-x_{ik,a}) \right] \, , \nonumber \\
            &\langle \boldsymbol{V}_{k}^{q_{a}q_{i}} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{g})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[(1-\epsilon)x_{ik,a} + 2\dfrac{1-x_{ik,a}}{x_{ik,a}} \right] \, ,
        \end{align}
        where $n_{s}(a)$ is the number of polarisations of particle $a$.
        For fermions $n_{s}(q) = n_{s}(\bar{q}) = 2$
        and for gluons $n_{s}(g) = d-2$. In the limit $\epsilon \rightarrow 0$
        all ratios of $n_{s}$ cancel to give unity.

    \subsection{Initial-initial dipoles}
        In the case of two initial-state partons, $a$ and $b$, there
        is an additional dipole contribution compared to the initial-final case.

        The dipole factorisation is given by
        \begin{equation}\label{eqn:II_factorisation}
            \begin{split}
                {}_{n+1} \langle 1, \ldots, n+1; a,b &| 1, \ldots, n+1;a,b \rangle_{n+1} = \\
                &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) \\
                + &\sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) + \ldots \, ,
            \end{split}
        \end{equation}
        where the new contribution is the dipole $\mathcal{D}^{ai,b}$ given by
        \begin{equation}\label{eqn:D_aib}
            \begin{split}
                \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) &= -\dfrac{1}{2p_{a}p_{i}} \dfrac{1}{x_{i,ab}} \\
                &{}_{n}\langle \tilde{1}, \ldots, \widetilde{n+1}; \widetilde{ai}, b | \dfrac{\boldsymbol{T}_{b}\cdot\boldsymbol{T}_{ai}}{\boldsymbol{T}^{2}_{ai}}\boldsymbol{V}^{ai,b} | \tilde{1}, \ldots, \widetilde{n+1}; \widetilde{ai}, b \rangle_{n} \, .
            \end{split}
        \end{equation}
        The initial-state parton $\widetilde{ai}$ is the emitter and the
        other initial-state parton $b$ is the spectator. Notice that the
        momenta not involved in the dipole term have been modified with only
        parton $b$ remaining unchanged. The momentum mapping for this case
        has $\tilde{p}_{ai}$ parallel with $p_{a}$ and also
        modifies the momenta for all other final-state momenta (even non-QCD particles) $k_{j}$:
        \begin{align}\label{eqn:II_mapping}
            \tilde{p}_{ai}^{\mu} &= x_{i,ab} \, p_{a}^{\mu} \, , \nonumber \\
            \tilde{k}_{j}^{\mu} &= k_{j}^{\mu} - \dfrac{2k_{j} \cdot (K+\widetilde{K})}{(K+\widetilde{K})^{2}}(K+\widetilde{K})^{\mu} + \dfrac{2k_{j} \cdot K}{K^{2}}\widetilde{K}^{\mu} \, , \\
            x_{i,ab} &= \dfrac{p_{a}p_{b} - p_{i}p_{a}-p_{i}p_{b}}{p_{a}p_{b}} \, . \nonumber
        \end{align}
        where $K$ and $\widetilde{K}$ are the total momenta
        of the dipole before and after the mapping, respectively. They are given by
        \begin{align}\label{eqn:II_mapping2}
            \begin{split}
            K^{\mu} &= p_{a}^{\mu} + p_{b}^{\mu} - p_{i}^{\mu} \, ,\\
            \widetilde{K}^{\mu} &= \tilde{p}_{ai}^{\mu} + p_{b}^{\mu} \, . \\
            \end{split}
        \end{align}
        This mapping conserves momentum and keeps mapped momenta on-shell
        \begin{align}\label{eqn:II_mapping_conditions}
            \begin{split}
            &p_{a}^{\mu} + p_{b}^{\mu} - p_{i}^{\mu} - \sum_{j} k_{j}^{\mu} = \tilde{p}_{ai}^{\mu} + p_{b}^{\mu} - \sum_{j} \tilde{k}_{j}^{\mu} = 0\, , \\
            &\tilde{p}_{ai}^{2} = \tilde{k}_{j}^{2} = 0 \, .
            \end{split}
        \end{align}
        The spin-averaged splitting functions are proportional
        to the Altarelli-Parisi splitting functions which we
        quote here for completeness
        \begin{align}\label{V_aib}
            \begin{split}
            &\langle \boldsymbol{V}^{q_{a}q_{i},b} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{g})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{1 + (1-x_{i,ab})^{2}}{x_{i,ab}}\right] \, , \\
            &\langle \boldsymbol{V}^{q_{a}g_{i},b} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{1 + x_{i,ab}^{2}}{1 - x_{i,ab}}\right] \, , \\
            &\langle \boldsymbol{V}^{g_{a}\bar{q}_{i},b} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[x_{i,ab}^{2} + (1-x_{i,ab})^{2}\right] \, , \\
            &\langle \boldsymbol{V}^{g_{a}g_{i},b} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{g})} \; 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{x_{i,ab}}{1-x_{i,ab}} + \dfrac{1-x_{i,ab}}{x_{i,ab}} + x_{i,ab}(1-x_{i,ab})\right] \, .
            \end{split}
        \end{align}

    \subsection{Master factorisation formula}
        The combination of all types of dipole contributions
        can be combined into a master dipole factorisation formula
        \begin{equation}\label{eqn:dipole_factorisation_master}
            \begin{split}
            {}_{n+1} \langle 1, &\ldots, n+1 | 1, \ldots, n+1 \rangle_{n+1} = \\
            &\sum_{i,j}\sum_{k \neq i, j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) \\
            + &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \ldots \, ,
            \end{split}
        \end{equation}
        where it is understood that for a given partonic process,
        the matrix element in the different soft and collinear limits will be
        reproduced by summing over all the relevant dipoles, meaning that
        the dipole functions can be used as a set of functional behaviours for
        approximating the matrix elements for arbitrary SM processes.
        This will be put into practice in Chapters \ref{chapter:fame1}
        and \ref{chapter:fame3}. The terms represented as ellipses in
        the dipole factorisation formulae are non-divergent
        in all of phase-space.
